Double White Dwarf Mergers with CASTRO\\ I. Methodology and Code Verification

### 2015-01-01 papers/wdmerger_I/paper.tex
Slight cleanup of hydrodynamics section, plus addition of the Euler equations.

@@

To study the white dwarf merger problem, we use the grid-based hydrodynamics code CASTRO \citep{castro}. CASTRO solves the Euler equations, along with the inclusion of optional modules for gravity, nuclear reactions and thermodynamics. We direct the reader to the original code paper for a description of CASTRO's approach to solving the equations of hydrodynamics. In this work, we report only on the changes we have made to the code since its original release, for the purpose of approaching this problem. CASTRO is based on the BoxLib adaptive-mesh refinement (AMR) framework (ref), which represents fluid data on a mesh where regions of interest have higher spatial resolution. CASTRO is highly parallel and is designed for large-scale use on modern supercomputers; see Section \ref{sec:Performance} for information on how CASTRO performs for our problem.

@@

The boundary conditions on the hyperbolic system are simply zero-gradient zones that allow material to flow directly out of the domain. Using AMR, we make the coarse grid very large, so as
to place the boundaries far from the region of interest. This ensures that any boundary effects do not pollute the inner region where the stars will eventually make contact.  We further
make the restriction that only level-0 (coarse) grids can touch the domain boundary.

@@

>>>>

To study the white dwarf merger problem, we use the grid-based hydrodynamics code CASTRO \citep{castro}. CASTRO solves the Euler equations, along with the inclusion of optional modules for gravity, nuclear reactions and thermodynamics. CASTRO is based on the BoxLib adaptive-mesh refinement (AMR) framework \citep{rendleman:2000}, which represents fluid data on a mesh where regions of interest have higher spatial resolution. CASTRO is highly parallel and is designed for large-scale use on modern supercomputers; see Section \ref{sec:Performance} for information on how CASTRO performs for our problem. The next few subsections describe our approach to each of the physics components used in this work. We direct the reader to the original code paper for a full description of CASTRO's approach to solving the equations of hydrodynamics. In this work, we report mainly on the changes we have made to the code since its original release, for the purpose of approaching this problem.

@@

The Euler equations for hydrodynamics (in the absence of source terms) in conservative form are:
\begin{align}
  \frac{\partial \rho}{\partial t} &= -\bm{\nabla} \cdot (\rho \mathbf{u}) \\
  \frac{\partial \rho \mathbf{u}}{\partial t} &= -\bm{\nabla} \cdot (\rho \mathbf{u}\mathbf{u}) - \bm{\nabla}p \\
  \frac{\partial \rho E}{\partial t} &= -\bm{\nabla}\cdot(\rho\mathbf{u}E + p\mathbf{u}).
\end{align}
Here $\rho$ is the mass density, $\mathbf{u}$ is the fluid velocity vector, $p$ is the pressure, and $E = \rho \mathbf{u}^2 / 2 + e$ is the total energy density where $e$ is the internal (thermal) energy density.

@@

@@

The boundary conditions on the hyperbolic system are simply zero-gradient zones that allow material to flow directly out of the domain. Using AMR, we make the coarse grid large enough that the boundaries are relatively far from the region of interest. This ensures that any boundary effects do not pollute the inner region where the stars will eventually make contact.  We further make the restriction that refined grids cannot reach the domain boundary.
